Categorical and gradient aspects of wordlikeness\thanks{Thanks to Gene Buckley, Charles Yang, and audiences at NELS 43 for helpful comments.}

### 2012-11-16 3.tex
Moving stuff around, mostly.

Much of the latter-day literature on phonotactic knowledge, while adopting the representational machinery of generative phonology, takes less interest in the question of what generalizations speakers internalize. In a seminal paper, \citet{McCarthy1988} uses a simple statistical technique to validate a generalization about Arabic root consonant co-occurrence facts originally due to \citet{Greenberg1950}. The assumption appears to be that static co-occurrence restrictions which have a statistically significant trace in the lexicon will also be reflected in psycholinguistic tasks such as wordlikeness judgements. While the Arabic restriction discussed by \citeauthor{McCarthy1988} was ultimately investigated experimentally \citep{Frisch2004}, this is the case for only a small minority of the studies inspired by \citeauthor{McCarthy1988}'s study; in most cases, phonotactic generalizations are inferred directly from lexical statistics \citep[e.g.,][]{Anttila2008a,Berkley1994b,Berkley1994a,Berkley2000,Brown2010,Buckley1997,Coetzee2008a,Elmedlaoui1995,Graff2011,MacEachern1999,Kinney2005,Kawahara2006,Martin2007,Martin2011,Mester1988,Miller-Ockhuizen2003,Padgett1992,Padgett1995,Pozdniakov2007,Yip1989}. In a study of consonant co-occurrence restrictions in Gitksan, \citet{Brown2010} claims that \emph{any} statistically significant pattern in the lexicon is one that is internalized by speakers:

@@

Finally, there are many cases of statistical generalizations with a known diachronic cause but lacking any synchronic reality. One of the earliest arguments of this type was put forth by Saussure (\citeyear[202f.]{CLG}) in a discussion of rhotacism, a diachronic process which changed Old Latin intervocalic \emph{s} to \emph{r}. As a result, Classical Latin intervocalic \emph{s} is rare. However, Saussure denies that this generalization has any status in Classical Latin.\footnote{See \citet{Gorman2012e} and citations therein for a full explication of Saussure's position on rhotacism, and a restatement of the phonological argument in generative terms.}

@@

\begin{table}
\centering
\begin{tabular}{l r r r r}
\toprule
          & \{ɪ, ɛ, æ, ʌ, ʊ\}\gap{}\# & \{i, e, ɑ, ɔ, u\}\gap{}\# & \% long & $p$-value \\
\midrule
\gap{}ʃ\# & 78                & 9                 & 8\%      & \multirow{2}{*}{0.026} \\
\gap{}s\# & 410               & 107               & 16\%     & \\
\bottomrule
\end{tabular}
\caption{Type frequencies of short and long vowels before word-final [ʃ] and [s] in a subset of the CMU pronunciation dictionary containing all words with token frequency greater than 1 per million words in the SUBTLEX-US frequency norms; the 9 words containing ending in a long vowel followed by ʃ are \emph{douche}, \emph{leash}, \emph{gosh}, \emph{josh}, \emph{posh}, \emph{squash}, \emph{unleash}, \emph{wash}, and \emph{woosh}.}
\label{ssh}
\end{table}

@@

Two less familiar notations are used in this chapter. First, directional application conditions are assumed, so as to derive the left-to-right, iterative properties of the harmony rules. Since it was first proposed by \citet{Johnson1972}, considerable evidence for directional application has been adduced (e.g., \citealt{A74}: chap.~9, \citealt{Gorman2012e}, \citealt{Howard1972}:65f., \citealt{Kavitskaya2008}, \citealt{Kaye1982}, \citealt{KK77}:189f., \citeyear{KK79}:319f., \citealt{Piggott1975}, \citealt{Sohn1971}; see \citealt{McCarthy2003b} and \citealt{Wolf2011b} for recent reviews.) Further, \citet{Johnson1972} and \citet{Kaplan1994} prove directional application and simultaneous application are equivalent in terms of formal learnability.

@@

\noindent
\citet[196]{Inkelas2001} propose instead that the trigger must contain a labial consonant but does not require a preceding \emph{a} vowel.

\begin{quote}
Vowel labialization following labials is not a synchronic alternation in Turkish, yet it (unlike \textsc{Labial Attraction} per se) \emph{is} a statistically supported tendency worthy of further research. \citep[][196]{Inkelas2001}
\end{quote}

@@

\begin{unlabeledexample}
$\displaystyle \gamma = \frac{c - d}{c + d}$
\end{unlabeledexample}

@@

\begin{table}
\centering
\begin{tabular}{lrlr|lrlr}
\toprule
\multicolumn{4}{c|}{Experiment 1} & \multicolumn{4}{c}{Experiment 2} \\
\multicolumn{2}{c}{aPu} & \multicolumn{2}{c|}{\textsc{disharmonic}} & \multicolumn{2}{c}{\textsc{harmonic}} & \multicolumn{2}{c}{\textsc{disharmonic}} \\
\midrule
{tamuz} & 3 & {tamız} & 16 & {pamuz} & 15 & {pamız} & 17 \\
{tafuz} & 3 & {tafız} & 17 & {tafuz} & 21 & {tafız} & 11 \\
{tavuz} & 9 & {tavız} & 4  & {mavuz} & 16 & {mavız} & 16 \\
{tapuz} & 7 & {tapız} & 9  & {tapuz} & 17 & {tapız} & 15 \\
{tabuz} & 5 & {tabız} & 12 & {tabuz} & 16 & {tabız} & 16 \\
\bottomrule
\end{tabular}
\caption{Effects of \textsc{Labial Attraction} on wordlikeness \citep[from][]{Zimmer1969}}
\label{law}
\end{table}

@@

\section{Conclusion}

>>>>

Much of the latter-day literature on phonotactic knowledge, while adopting the representational machinery of generative phonology, takes less interest in the question of what generalizations speakers internalize. In a seminal paper, \citet{McCarthy1988} uses a simple statistical technique to validate a generalization about Arabic root consonant co-occurrence facts originally due to \citet{Greenberg1950}. The assumption appears to be that static co-occurrence restrictions which have a statistically significant trace in the lexicon will also be reflected in psycholinguistic tasks such as wordlikeness judgements. While the Arabic restriction discussed by \citeauthor{McCarthy1988} was ultimately investigated experimentally \citep{Frisch2004}, this is the case for only a small minority of the studies inspired by \citeauthor{McCarthy1988}'s study; in most cases, phonotactic generalizations are inferred directly from lexical statistics \citep[e.g.,][]{Anttila2008a,Berkley1994b,Berkley1994a,Berkley2000,Brown2010,Buckley1997,Coetzee2008a,Elmedlaoui1995,GraffInPress,MacEachern1999,Kinney2005,Kawahara2006,Martin2007,Martin2011,Mester1988,Miller-Ockhuizen2003,Padgett1992,Padgett1995,Pozdniakov2007,Yip1989}. In a study of consonant co-occurrence restrictions in Gitksan, \citet{Brown2010} claims that \emph{any} statistically significant pattern in the lexicon is one that is internalized by speakers:

@@

Finally, there are many cases of statistical generalizations with a known diachronic cause but lacking any synchronic reality. One of the earliest arguments of this type was put forth by Saussure (\citeyear[202f.]{CLG}) in a discussion of rhotacism, a diachronic process which changed Old Latin intervocalic \emph{s} to \emph{r}. As a result, Classical Latin intervocalic \emph{s} is rare. However, Saussure denies that this generalization has any status in Classical Latin.\footnote{See \citet{GormanInPressc} and citations therein for a full explication of Saussure's position on rhotacism, and a restatement of the phonological argument in generative terms.}

@@

\begin{table}
\centering
\begin{tabular}{l r r r r}
\toprule
          & \{ɪ, ɛ, æ, ʌ, ʊ\}\gap{}\# & \{i, e, ɑ, ɔ, u\}\gap{}\# & \% long & $p$-value \\
\midrule
\gap{}ʃ\# & 78                & 9                 & 8\%      & \multirow{2}{*}{.026} \\
\gap{}s\# & 410               & 107               & 16\%     & \\
\bottomrule
\end{tabular}
\caption{Type frequencies of short and long vowels before word-final [ʃ] and [s] in a subset of the CMU pronunciation dictionary containing all words with token frequency greater than 1 per million words in the SUBTLEX-US frequency norms; the 9 words containing ending in a long vowel followed by ʃ are \emph{douche}, \emph{leash}, \emph{gosh}, \emph{josh}, \emph{posh}, \emph{squash}, \emph{unleash}, \emph{wash}, and \emph{woosh}.}
\label{ssh}
\end{table}

@@

Two less familiar notations are used in this chapter. First, directional application conditions are assumed, so as to derive the left-to-right, iterative properties of the harmony rules. Since it was first proposed by \citet{Johnson1972}, considerable evidence for directional application has been adduced (e.g., \citealt{A74}: chap.~9, \citealt{GormanInPressc}, \citealt{Howard1972}:65f., \citealt{Kavitskaya2008}, \citealt{Kaye1982}, \citealt{KK77}:189f., \citeyear{KK79}:319f., \citealt{Piggott1975}, \citealt{Sohn1971}; see \citealt{McCarthy2003b} and \citealt{Wolf2011b} for recent reviews.) Further, \citet{Johnson1972} and \citet{Kaplan1994} prove directional application and simultaneous application are equivalent in terms of formal learnability.

@@

\noindent
\citet{Inkelas2001} propose instead that the trigger must contain a labial consonant but does not require a preceding \emph{a} vowel.

\begin{quote}
Vowel labialization following labials is not a synchronic alternation in Turkish, yet it (unlike \textsc{Labial Attraction} per se) \emph{is} a statistically supported tendency worthy of further research. \citep[196]{Inkelas2001}
\end{quote}

@@

\begin{equation*}
\displaystyle \gamma = \frac{c - d}{c + d}
\end{equation*}

@@

\begin{table}
\centering
\begin{tabular}{lrlr|lrlr}
\toprule
\multicolumn{4}{c|}{Experiment 1} & \multicolumn{4}{c}{Experiment 2} \\
\multicolumn{2}{c}{aPu} & \multicolumn{2}{c|}{aPı} & \multicolumn{2}{c}{aPu} & \multicolumn{2}{c}{aPı} \\
\midrule
{tamuz} & 3 & {tamız} & 16 & {pamuz} & 15 & {pamız} & 17 \\
{tafuz} & 3 & {tafız} & 17 & {tafuz} & 21 & {tafız} & 11 \\
{tavuz} & 9 & {tavız} & 4  & {mavuz} & 16 & {mavız} & 16 \\
{tapuz} & 7 & {tapız} & 9  & {tapuz} & 17 & {tapız} & 15 \\
{tabuz} & 5 & {tabız} & 12 & {tabuz} & 16 & {tabız} & 16 \\
\bottomrule
\end{tabular}
\caption{Effects of \textsc{Labial Attraction} on wordlikeness \citep[from][]{Zimmer1969}}
\label{law}
\end{table}

@@

\section{Conclusions}

### 2012-11-16 4.tex
Moving stuff around, mostly.

English syllable contact clusters have been used, by \citeauthor{Pierrehumbert1994} and others, to argue for the necessity of admitting static phonotactic constraints, and to illustrate proposals for the architecture of the phonotactic system. 
However, there are a number of reasons to reconsider the facts of this domain.

@@

\citeauthor{Pierrehumbert1994} does not fully consider the effects of phonological processes targeting word-medial clusters in English, and consequently some of the static constraints she proposes may be related to English morphophonemics. 
For instance, she writes that ``nasal-stop sequences agree in labiality'' (175), and attributes this the status of a static constraint. 
However, this generalization is just a narrower form of the one derived by \textsc{Nasal Place Assimilation} (see \S\ref{npa}), and which requires no further explanation. 
Other derived constraints go unnoticed: for instance, she does not discuss the highly reliable tendency of obstruent-obstruent clusters to agree in voicing (see \S\ref{ova}). 
The evaluation below considers the role of both types of constraints on the English lexicon on an equal footing.

@@

Many components of \citeauthor{Pierrehumbert1994}'s study cannot be directly replicated. \citeauthor{Pierrehumbert1994} focuses her attention on those words she judges to be ``morphologically simple'' and ``reasonably familiar'', but this list has neither been published nor circulated.
It is not clear that the judgements (as well as cognitive limitations) of concerned parties should be granted evidential status in the first place, given the potential for implicit bias; 
\citet{Labov1975}, for instance
argues for what he calls the \emph{Experimenter Principle}, 
that the intuitions of all those familiar with the formal matters should be excluded.
Otherwise-unmotivated morphological junctures are sometimes posited simply to preserve phonological or phonotactic generalizations; 
%The same practice is sometimes applied to preserve phonological generalizations; for instance, otherwise unmotivated morphological boundaries are used in \emph{SPE} to simplify principles of stress assignment. 
this is done by \citet{SPE}, for instance, to simplify principles of English stress assignment, and \citet[546]{Rice2009d} analyses many words in Slave as compounds simply because they contain consonant clusters that rarely occur morph-internally. 
%There is some evidence that infants \citep{Mattys2001b} and adults \citep{Brown1956,Hay2004a,McQueen1998b,Norris1997} use this heuristic to segment fluent speech in experimental settings. 
Applied indiscriminately, however, such a heuristic trivializes both morphological segmentation and phonotactic generalization.
For these reasons, the wordlist used in this study is derived from a publicly available database, and no experimenter intuitions are used for establishing morphological boundaries.

@@

Following \citet[ chap.~8]{Duanmu2009} and \citet[ chap.~3]{Hammond1999a}, 
a wordlist is generated using the English portion of the CELEX database \citep{CELEX}. 
Only words coded as ``monomorphemic'' are used, and all words coded as non-native are excluded.
This stringent criterion for inclusion eliminates many of the exceptions noted by \citeauthor{Duanmu2009} or \citeauthor{Hammond1999a} in their studies. For instance, nearly all the exceptions to \textsc{Obstruent Voice Assimilation} (see \S\ref{ova} below) noted by \citet[74]{Hammond1999a} are 
either complex words (e.g., \emph{jurisdiction}, \emph{madcap} \emph{tadpole}, \emph{scapegoat}, \emph{magpie}) or non-native (e.g., \emph{vodka}, \emph{smorgasbord}).

Unlike prior studies, however, words which appear to consist of a Latinate prefix and a bound stem (e.g., \emph{inspect}, \emph{excrete}) are also excluded from the sample, as there is broad formal and experimental support for the claim that these words are complex.
This assumption allows for numerous simplifications to accounts of the phonological shape of these forms. 
For instance, \citet[11f.]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation.

@@

\noindent
\citeauthor{Aronoff1976} takes this to be evidence that \emph{adhere} and \emph{cohere}, for instance, share a bound stem.
There is also an interaction between Latinate prefixes on verbs and the makeup of the verbal complement. 
Latinate verbs do not generally allow ditransitive, verb participle, or adjectival resultative constructions, all acceptable with similar Anglo-Saxon verbs \citep{Gropen1989,Harley2009}.
%\citep[see also][]{Gropen1989,Coppock2008}.

\begin{example}[Latinate verbs and small clauses] 
\label{harley}
\begin{tabular}{l l l l@{} l}
a. & {show him the painting} & \alt{} & * & {exhibit him the painting} \\
b. & {drink himself stupid}  & \alt{} & * & {imbibe himself stupid}    \\
c. & {show it off}           & \alt{} & * & {exhibit it off}           \\
\end{tabular}
\end{example}

@@

This results in a list of 6,619 simplex words. 
The full set of clusters and their frequencies are given in Appendix \ref{clusters}.
The CELEX transcriptions of these words were then syllabified and phonologized using a technique described in Appendix \ref{syllabification}. 
This produces 23 unique medial codas and 40 unique medial onsets. 
Of the 920 ($= 21 \times 40$) medial clusters that would result from free combination, 174 (19\%) are attested.

@@

So as to account for the 81\% of ``possible'' but unattested clusters, \citet{Pierrehumbert1994} proposes three constraints on English medial clusters. These constraints are static, in the sense that they have no analogues in the morphophonemics of English.

\paragraph{Dorsal-labial clusters} 
\citet[173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars'', noting that velar-velar clusters are independently excluded by constraints on geminates (see \S?). 
Two-consonant velar-labial clusters are found in words like \emph{a}[k.m]\emph{e}, \emph{ru}[ɡ.b]\emph{y}, or \emph{pi}[ɡ.m]\emph{ent} occur. 
Table \ref{dltab} shows that velar-labial clusters are somewhat less likely to occur than  are velar-coronal clusters (e.g., \emph{ve}[k.t]\emph{or}), but such underrepresentation is not unlikely to occur by chance ($p = .106$).

@@

\paragraph{Degemination} The final alternation found in English medial clusters is the simplification of geminates which is characteristic of ``level I'' morphology, and is found in the irregular /-t/ past tense (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]) and in \emph{-ly} deadjectival derivatives (e.g., \emph{norma}[l]\emph{y}, cf. \emph{calm}[l]\emph{y}), and Latinate prefix allomorphy (\emph{SPE}:148; \citealt[102]{Borowsky1986}). \textsc{Degemination} is formalized here as a rule deleting the first of two segments agreeing on all feature values (except voice, possibly).

@@

\paragraph{Summary} All three of \emph{SPE} rules targeting medial clusters have a robust effect in constraining the inventory of possible word-medial syllable clusters; possible clusters which are surface exceptions to these three rules are much less likely to be attested than those which conform to them.

@@

Current computational models of phonotactic knowledge ``rate'' possible clusters by producing a numerical value for each cluster. 
To apply these models to predicting the cluster inventory, these 
these numerical values must be transformed into a categorical prediction, of either attestation or non-attestation.
This is accomplished with a soft-margin support vector machine \citep{Cortes1995} with a linear kernel, which optimally splits the ratings distribution into a set of attested and unattested clusters.
This classifier is does not correspond to any component of a cognitively plausible model of phonotactic learning: it simply represents an upper bound for predicting the cluster inventory from positive data.

@@

It can be shown that the large number of possible-but-unattested clusters is a logical necessity, given the sparse nature of the coda and onset frequencies. 
In Figure \ref{cao}, the rank and type frequency (i.e., frequency in the lexicon) of medial codas and onsets in the lexical sample described below are graphed in log-log space. 
Both codas and onsets show a linear relationship between log rank and log frequency characteristic of Zipfian distributions (see Appendix \ref{zr}). 
As a result, an enormous lexical sample would be needed to realize clusters consistent with the \textsc{Medial Cluster Law} even if there were no constraints on the combination of medial codas and onsets in English.
To illustrate this fact, a simulations is used to create new ``samples'' of the same size as the lexical sample used here. The following procedure is repeated so as to generate each observation in the simulated sample.

>>>>

English syllable contact clusters have been used, by \citeauthor{Pierrehumbert1994} and others, to argue for the necessity of admitting static phonotactic constraints, and to illustrate proposals for the architecture of the phonotactic system. 
However, there are a number of reasons to reconsider this domain.

@@

\citeauthor{Pierrehumbert1994} does not fully consider the effects of phonological processes targeting word-medial clusters in English, and consequently some of the static constraints she proposes may be related to English morphophonemics. 
For instance, she writes that ``nasal-stop sequences agree in labiality'' (175), and attributes this the status of a static constraint. 
However, this generalization is just a narrower form of the one derived by \textsc{Nasal Place Assimilation} (see \S\ref{npa}).
Other derived constraints go unnoticed: for instance, she does not discuss the highly reliable tendency of obstruent-obstruent clusters to agree in voicing (see \S\ref{ova}). 
The evaluation below considers the role of both types of constraints on the English lexicon on an equal footing.

@@

Many components of \citeauthor{Pierrehumbert1994}'s study cannot be directly replicated. \citeauthor{Pierrehumbert1994} focuses her attention on those words she judges to be ``morphologically simple'' and ``reasonably familiar'', but this list has neither been published nor circulated.
It is not clear that the judgements (as well as cognitive limitations) of concerned parties should be granted evidential status in the first place, given the potential for implicit bias; 
\citet{Labov1975} for instance
argues for what he calls the \emph{Experimenter Principle}, 
that the intuitions of all those familiar with the formal matters should be excluded.
Otherwise-unmotivated morphological junctures are sometimes posited simply to preserve phonological or phonotactic generalizations; 
%The same practice is sometimes applied to preserve phonological generalizations; for instance, otherwise unmotivated morphological boundaries are used in \emph{SPE} to simplify principles of stress assignment. 
this is done by \citet{SPE}, for instance, to simplify principles of English stress assignment, and \citet[546]{Rice2009d} analyses many words in Slave as compounds simply because they contain consonant clusters that rarely occur morph-internally. 
%There is some evidence that infants \citep{Mattys2001b} and adults \citep{Brown1956,Hay2004a,McQueen1998b,Norris1997} use this heuristic to segment fluent speech in experimental settings. 
Applied indiscriminately, however, such a heuristic trivializes both morphological segmentation and phonotactic generalization.
For these reasons, the wordlist used in this study is derived from a publicly available database, and no experimenter intuitions are used for establishing morphological boundaries.

@@

Following \citet[ chap.~8]{Duanmu2009} and \citet[ chap.~3]{Hammond1999a}, 
a wordlist is generated using the English portion of the CELEX database \citep{CELEX}. 
Only words coded as ``monomorphemic'' are used, and all words coded as non-native are excluded.
This stringent criterion for inclusion eliminates many of the exceptions noted by \citeauthor{Duanmu2009} or \citeauthor{Hammond1999a} in their studies. For instance, nearly all the exceptions to \textsc{Obstruent Voice Assimilation} (see \S\ref{ova} below) noted by \citet[74]{Hammond1999a} are 
either complex words (e.g., \emph{jurisdiction}, \emph{madcap}, \emph{tadpole}, \emph{scapegoat}, \emph{magpie}) or non-native (e.g., \emph{vodka}, \emph{smorgasbord}).

Unlike prior studies, words which appear to consist of a Latinate prefix and a bound stem (e.g., \emph{inspect}, \emph{excrete}) are also excluded from the sample, as there is broad formal and experimental support for the claim that these words are complex.
This assumption allows for numerous simplifications to accounts of the phonological shape of these forms. 
For instance, \citet[11f.]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation.

@@

\noindent
\citeauthor{Aronoff1976} takes this to be evidence that \emph{adhere} and \emph{cohere}, for instance, share a bound stem.
There is also an interaction between Latinate prefixes on verbs and the makeup of the verbal complement. 
Latinate verbs do not generally allow ditransitive, verb participle, or adjectival resultative constructions, all of which are acceptable with similar Anglo-Saxon verbs \citep{Gropen1989,Harley2009}.
%\citep[see also][]{Gropen1989,Coppock2008}.

\begin{example}[Latinate verbs and small clauses] 
\label{harley}
\begin{tabular}{l l l l@{} l}
a. & {show him the painting} & \alt{} & * & {exhibit him the painting} \\
b. & {drink himself stupid}  & \alt{} & * & {imbibe himself stupid}    \\
%c. & {show it off}           & \alt{} & * & {exhibit it off}           \\
c. & {break it off}           & \alt{} & * & {terminate it off}        \\
\end{tabular}
\end{example}

@@

The filters discussed above result in a list of 6,619 simplex words.
The full set of clusters and their frequencies are given in Appendix \ref{clusters}.
The CELEX transcriptions of these words were then syllabified and phonologized using a technique described in Appendix \ref{syllabification}. 
This produces 23 unique medial codas and 40 unique medial onsets. 
Of the 920 ($= 21 \times 40$) medial clusters that would result from free combination, 174 (19\%) are attested.

@@

So as to account for the 81\% of ``possible'' but unattested clusters, \citet{Pierrehumbert1994} proposes three constraints on English medial clusters. These constraints are static, in that they have no analogues in the morphophonemics of English.

\paragraph{Dorsal-labial clusters} 
\citet[173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars'', noting that velar-velar clusters are independently excluded by constraints on geminates (see \S\ref{degemx}). 
Two-consonant velar-labial clusters are found in words like \emph{a}[k.m]\emph{e}, \emph{ru}[ɡ.b]\emph{y}, or \emph{pi}[ɡ.m]\emph{ent} occur. 
Table \ref{dltab} shows that velar-labial clusters are somewhat less likely to occur than  are velar-coronal clusters (e.g., \emph{ve}[k.t]\emph{or}), but such underrepresentation is not unlikely to occur by chance ($p = .106$).

@@

\paragraph{Degemination} 
The final alternation found in English medial clusters is the simplification of geminates which is characteristic of ``level I'' morphology, and is found in the irregular /-t/ past tense (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]) and in \emph{-ly} deadjectival derivatives (e.g., \emph{norma}[l]\emph{y}, cf. \emph{calm}[l]\emph{y}), and Latinate prefix allomorphy (\emph{SPE}:148; \citealt[102]{Borowsky1986}). \textsc{Degemination} is formalized here as a rule deleting the first of two segments agreeing on all feature values (except voice, possibly).
\label{degemx}

@@

\paragraph{Summary} All three of the \emph{SPE} rules targeting medial clusters have a robust effect in constraining the inventory of possible word-medial syllable clusters; possible clusters which are surface exceptions to these three rules are much less likely to be attested than those which conform to them.

@@

Current computational models of phonotactic knowledge ``rate'' possible clusters by producing a numerical value for each cluster. 
To apply these models to predicting the cluster inventory, these numerical values must be transformed into a categorical prediction of either attestation or non-attestation.
This is accomplished with a soft-margin support vector machine \citep{Cortes1995} with a linear kernel, which optimally splits the ratings distribution into a set of attested and unattested clusters.
This classifier does not correspond to any component of a cognitively plausible model of phonotactic learning: it simply represents an upper bound for predicting the cluster inventory from positive data.

@@

It can be shown that the large number of possible-but-unattested clusters is a logical necessity, given the sparse nature of the coda and onset frequencies. 
In Figure \ref{cao}, the rank and type frequency (i.e., frequency in the lexicon) of medial codas and onsets in the lexical sample described below are graphed in log-log space. 
Both codas and onsets show a linear relationship between log rank and log frequency characteristic of Zipfian distributions (see Appendix \ref{zr}). 
As a result, an enormous lexical sample would be needed to realize clusters consistent with the \textsc{Medial Cluster Law} even if there were no constraints on the combination of medial codas and onsets in English.
To illustrate this fact, a simulation is used to create new ``samples'' of the same size as the lexical sample used here. The following procedure is repeated so as to generate each observation in the simulated sample.

### 2012-11-16 abs.tex
Moving stuff around, mostly.

This dissertation argues for that phonotactic knowledge is primarily derived not from static generalizations but rather from phonological alternations and prosodic parsing. Chapter \ref{intro} outlines the proposal, historical precursors, and predictions. 
%especially with regards to order of acquisition.

Chapter \ref{gradience} considers evidence from wordlikeness rating tasks. It is argued that intermediate well-formedness ratings are obtained whether or not the categories in question are graded. A primitive categorical model of wordlikeness using prosodic representations is outlined, and shown to predict English speakers' wordlikeness judgements at least as accurately as state-of-the-art gradient wellformedness models. Once categorical effects are controlled for, gradient models are uncorrelated with well-formedness ratings.

>>>>

This dissertation outlines a program for the theory of phonotactics---the theory of speakers' knowledge of possible and impossible (or likely and unlikely) words---minimizing undue \emph{duplication} (in the sense of \citealt{KK77}:136f.) of phonotactics and phonology, and argues that the alternative view of phonotactics as stochastic, and of phonotactic learning as probabilistic inference, is not capable of accounting for the facts of this domain.
Chapter \ref{intro} outlines the proposal, precursors, and predictions. 
%especially with regards to order of acquisition.

Chapter \ref{gradience} considers evidence from wordlikeness rating tasks. It is argued that intermediate well-formedness ratings are obtained whether or not the categories in question are graded. A primitive categorical model of wordlikeness using prosodic representations is outlined, and shown to predict English speakers' wordlikeness judgements as accurately as state-of-the-art gradient wellformedness models. Once categorical effects are controlled for, gradient models are uncorrelated with well-formedness ratings.

### 2012-11-16 ms2.tex
Moving stuff around, mostly.

\abstract{
Gradient wordlikeness judgements do not necessarily imply that a gradient well-formedness system underlies them; gradient judgements may be an artifact of gradient rating tasks.
Dubious architectural assumptions are needed for speakers to report gradient well-formedness judgements.
Simple baselines better account for gradient well-formedness judgements than state-of-the-art computational models of gradient phonotactic knowledge.
}
\section{Introduction}
\input{2}
\appendix
\renewcommand{\arraystretch}{0.25}

@@

Initially, between two vowels, or finally, \emph{r} is consonantal. Before another consonant, however, \emph{r} has been lost in Received Pronunciation. Even in \emph{r}-ful dialects, though, post-vocalic non-onset \emph{r} patterns with vowels, not coda consonants. Before non-onset \emph{r} many vowel contrasts are suspended (e.g., \citealp[269f.]{Fudge1969}, \citealp[][255]{Harris1994}): compare American English \emph{fern}/\emph{fir}/\emph{fur} to \emph{pet}/\emph{pit}/\emph{putt}. In this position, \emph{r} is the only consonant which permits variable glottalization of a following /t/ in \emph{r}-ful British dialects \citep[258]{Harris1994}, and the only consonant which does does not trigger variable deletion of a following word-final /t, d/ in American dialects \citep[8]{Guy1980}. This is shown in (\ref{tglottalization}--\ref{tddeletion}) below.

>>>>

\abstract{
Gradient wordlikeness judgements do not necessarily imply that there is a gradient well-formedness system underlying them; gradient judgements may be an artifact of gradient rating tasks.
Dubious architectural assumptions are needed for speakers to report gradient well-formedness judgements.
Simple baselines better account for gradient well-formedness judgements than state-of-the-art computational models of gradient phonotactic knowledge.
}
\section{Introduction}
\input{2}
\appendix
\renewcommand{\arraystretch}{0.25}

@@

Initially, between two vowels, or finally, \emph{r} is consonantal. Before another consonant, however, \emph{r} has been lost in Received Pronunciation. Even in \emph{r}-ful dialects, though, post-vocalic non-onset \emph{r} patterns with vowels, not coda consonants. Before non-onset \emph{r} many vowel contrasts are suspended (e.g., \citealp[269f.]{Fudge1969}, \citealp[][255]{Harris1994}): compare American English \emph{fern}/\emph{fir}/\emph{fur} to \emph{pet}/\emph{pit}/\emph{putt}. In this position, \emph{r} is the only consonant which permits variable glottalization of a following /t/ in \emph{r}-ful British dialects \citep[258]{Harris1994}, and the only consonant which does not trigger variable deletion of a following word-final /t, d/ in American dialects \citep[8]{Guy1980}. This is shown in (\ref{tglottalization}--\ref{tddeletion}) below.
